Blue Hunters: Bluetooth RSSI Locator Robots

### 2018-02-06 article/article.tex
Bruce fixes + refs

\subsubsection{Chassis}

% Possible subsections?
% - building robot
% - signal detection
% - navigation

@@

\begin{figure}
  \centering
  \subfloat[Top view]{
    \includegraphics[width=0.4\textwidth]{top_view.jpg}
    \label{fig:robottop}
  }
  \subfloat[Side view]{
    \includegraphics[width=0.4\textwidth]{side_view.jpg}
    \label{fig:robotside}
  }
  \\
  \subfloat[Front view]{
    \includegraphics[width=0.4\textwidth]{front_view.jpg}
    \label{fig:robotfront}
  }
  \subfloat[Back view]{
    \includegraphics[width=0.4\textwidth]{inside_view.jpg}
    \label{fig:robotback}
  }
  \caption{Multiple views of the completed robots}
  \label{fig:robotviews}
\end{figure}

\hypertarget{electronics}{%
\subsubsection{Electronics}\label{electronics}}

Each robot consisted of 4 main components:
\begin{enumerate}
  \item PIC32MC250F128B microcontroller.
  \item MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer.
  \item HM-10 BLE module with a Texas Instruments CC2541 chip.
  \item Two 9 gram FS90R micro continuous rotation servos.
\end{enumerate}

@@

\paragraph{Bluetooth modules}

The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao,
\footnote{Jnhuamao is the company which makes the HM-10 module, along with other Bluetooth modules. Their English website is: \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.}
and did not come with genuine Jnhuamao firmware.
However, they did have genuine TI CC2541 chips.
We realized they were fakes when they did not behave according to
the Jnhuamao data sheet.
\footnote{The original data sheet comes from \cite{jnhuamaodatasheet}.
A PDF is available at \cite{jnhuamaomit}, which may be out of date, but is eaiser to get.}
Luckily, the hardware on the fake chips is the same as that of the genuine chips, minus an external crystal, and the genuine firmware checks for the presence of the crystal, and works even without
it. \cite{crystal}
As such, we reprogrammed the chips with the genuine firmware according
to an Arduino Fourm post \cite{crystal}:

@@

\subsection{Software}

The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
compiler and the PIC 32 Legacy Peripheral Library (plib). \footnote{
The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}}

\subsubsection{BLE}

@@

@@

\paragraph{IMU}

The PIC communicates with the IMU via I2C. The QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module has 2 dies:
one contains the 3-axis gyroscope and 3-axis accelerometer, which were not
used in this project, and the other die is the AK8963 3-axis
magnetometer (compass). \cite{ak8963cdatasheet}

The compass is connected to the rest of the MPU module via an auxiliary I2C bus.
In order to make the compass an accessible slave on the I2C bus, pass-through mode to be enabled on the IMU.
\footnote{We based our IMU code off of another ECE 4760 project, the Self-Balancing Robot. \cite{selfbalancingrobot} Their \texttt{i2c\_helper.h} was very helpful.}
The AK8963 has several modes of operation, and the chip must be set to power-down mode before switching to other modes.
We read compass values with single measurement mode, as in figure \ref{fig:imu_single_measurement}:

\begin{enumerate}
\item
  Set the compass to single measurement mode in 14 bit resolution.
\item
  Read the 6 data registers (X low, X high, Y low, Y high, Z low, Z
  high)
\item
  Read the Status 2 register to check for magnetic sensor overflow.
  Without reading this register, the read is not considered complete and
  further reads will fail.
\item
  Wait; if the IMU is read too frequently, it will not have enough time
  to take measurements.
\end{enumerate}

@@

% \paragraph{Servos}

% \paragraph{Gradient descent}

@@

This project was an interesting exploration into short-range distance
determination using Bluetooth, a generally unconventional approach.
We knew that Bluetooth RSSI would be noisy, mostly due to multipath interference.
The robots worked reliably when they stayed within roughly 1 meter of the beacon.
After this, they entered the land of shallow gradients:
the signal strength from the beacon (already noisy) would not change very
much, and often only due to noise.
They normally could never recover from this.

>>>>

\subsubsection{Building the Robot}

@@

% TODO: move around to be more top-down
Each robot consisted of 4 main components: a PIC32MC250F128B microcontroller, a MPU-9250 IMU with a 3 axis gryoscope, accelerometer, and magnetometer, a HM-10 BLE module with a Texas Instruments CC2541 chip, and two 9 gram FS90R micro continuous rotation servos.

@@

The robots were programmed in C, using MPLAB X IDE v4.0, the XC32 v1.4
compiler and the PIC 32 Legacy Peripheral Library (plib).
The full source code is available in online: \url{https://github.com/orangeturtle739/bluehunters/tree/master/ble.X}

\subsubsection{Signal Detection: Bluetooth Modules}

% Jnhuamao is the company which makes the HM-10 module, along with other Bluetooth modules. Their English website is: \url{http://www.jnhuamao.cn/bluetooth.asp?id=1}.

The HM-10 bluetooth modules we bought off Ebay were fakes: they were not made by Jnhuamao , and did not come with genuine Jnhuamao firmware.
However, they did have genuine TI CC2541 chips.
We realized they were fakes when they did not behave according to
the Jnhuamao data sheet at \cite{jnhuamaodatasheet}, 
with a PDF available at \cite{jnhuamaomit}, which may be out of date, but is easier to get.
Luckily, the hardware on the fake chips is the same as that of the genuine chips, minus an external crystal, and the genuine firmware checks for the presence of the crystal, and works even without
it. \cite{crystal}
As such, we reprogrammed the chips with the genuine firmware according
to an Arduino Fourm post \cite{crystal}:

@@

@@

% TODO: Tablify

@@

\subsubsection{Navigation}

After measuring the proximity to the beacon, the robots would need to move towards it. The robot made use of the magnetometer (compass) inside the IMU to orient itself, and the gradient descent algorithm to decide the direction of its next movement.

For the IMU, we used the QFN MPU-9250 \cite{mpu9250datasheet} \cite{mpu9250regmap} module. It has 2 dies: one is the AK8963 3-axis magnetometer \cite{ak8963cdatasheet}, and the other contains the 3-axis gyroscope and 3-axis accelerometer, which were not used in this project.

The microcontroller communicates with the IMU via I2C, and the compass is connected to the rest of the MPU module via an auxiliary I2C bus. For communication between the microcontroller and the AK8963's main die, we based our work off of basic I2C functions from another ECE 4760 project, the Self-Balancing Robot \cite{selfbalancingrobot}. In particular, their \texttt{i2c\_helper.h} was very helpful.

For communication with the actual magnetometer, we needed to configure the IMU to set the compass must as a slave on the I2C bus. For this, it was necessary to enable pass-through mode on the IMU during its configuration.

The AK8963 has several modes of operation, and the chip must be set to power-down mode before switching to other modes.
We read compass values with single measurement mode, as in figure \ref{fig:imu_single_measurement}.

A single compass read involves setting the compass to single measurement mode in 14 bit resolution, reading the 6 data registers (X low, X high, Y low, Y high, Z low, Z high), reading the Status 2 register to check for magnetic sensor overflow (without reading this register, the read is not considered complete and further reads will fail), and finally waiting to ensure that the IMU is not read too frequently, in which case it will not have enough time to take measurements.

@@

@@

This project was an interesting exploration into short-range distance
determination using Bluetooth, a generally unconventional approach.
We knew that Bluetooth RSSI would be noisy, mostly due to multipath interference and the presence of multiple transmitters in testing environments.
The robots worked reliably when they stayed within roughly 1 meter of the beacon.
After this, they entered the land of shallow gradients:
the signal strength from the beacon (already noisy) would not change very
much, and often only due to noise.
They normally could never recover from this.
